\vspace{1.0cm} \LARGE \bf Revealing the competitive dynamics of influenza viruses through evolutionary cartography

### 2012-11-29 flux.tex
Include section on virus/serum effects.

>>>>

\subsection*{Incorporating virus and serum effects}

The preceding model represents immunological distance as a drop in titer against the most reactive comparison for a particular antiserum.  
Here, we relax this assumption and treat the maximum reactivity of a sera as a random variable.
In this case, $H_{ij}$ still follows
\begin{equation} 
	H_{ij} \sim \mbox{Normal}( S_j - \delta_{ij}, \mdssd^2 ),
\end{equation}
but the vector of `serum effects' $\mathbf{S}$ is estimated rather than fixed.
We assume that $S_j$ values are hierarchically distributed according to a normal distribution.
The mean and variance of this distribution is taken from the empirical mean and empirical variance of the set of maximum titers across sera $\{ \max ( H_{1j},\ldots,H_{\vn j} ) : j = 1,\ldots,\sn \}$.
This formulation assumes that particular sera are more reactive in general than other sera.

Additionally, we follow the same logic and assume that particular virus isolates are more reactive in general than other virus isolates.
With virus reactivity included, $H_{ij}$ follows
\begin{equation} 
	H_{ij} \sim \mbox{Normal}\left(  \frac{V_i+S_j}{2} - \delta_{ij}, \mdssd^2 \right),
\end{equation}
and the vector of `virus effects' $V_i$ for $i = 1,\ldots, \vn$ is estimated in an analogous hierarchical fashion, with $\mathbf{V}$ normally distributed with mean and variance equal to the empirical mean and variance of the set of maximum titers across viruses $\{ \max ( H_{i1},\ldots,H_{i \sn} ) : i = 1,\ldots,\vn \}$.

With these effects included, Metropolis-Hastings proposals additionally incorporate moves to individual serum effects $S_j$ and virus effects $V_i$.
